Draft -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- Draft

### 2024-11-12 Reply_to_reviewers.qmd

---
title: "Replies to reviewers of the 1st round of Remote Sensing"
author: "Simon Oiry"
format: docx
editor: visual
---

# Reviewer 1

## Reviewer's comments

Overall, this paper presents a solidly researched and theoretically grounded study, offering insights into the types of seagrass bed vegetation and the importance of seagrass bed monitoring. It is a worthwhile document to read. However, there are issues related to uniqueness and experimental organization that, if addressed, could enhance its academic quality:

1.  Please highlight more clearly the innovations, distinctive features, or unique content of this paper. The work conducted in this paper is indeed sufficiently rigorous, but the unique content or scientific contributions of this work are not yet prominent enough. Since the equipment and methods used in this paper are relatively mature, the distinctive contributions of the current work are not sufficient to support its scientific contributions. Please elaborate and summarize in this regard, and if possible, consider dedicating a separate chapter for a detailed explanation.

2.  There seem to be issues with the classification results presented in Figures 5 to 8. For instance, in Figure 5, the Magnoliopsida category includes two different spectral representations of ground objects (appearing as green and brown in the RGB image). Similarly, in Figure 8, Magnoliopsida encompasses multiple spectral representations of ground objects (displayed in various colors in the RGB image). Please carefully organize and analyze the experimental results. The current figures have some Conflicts with the accuracy table and confusion matrix.

## Reply

1.  Thank you for this valuable suggestion. We appreciate your feedback on the need to further highlight the unique contributions and innovations of our study. In response, we have revised the final paragraph of the Introduction (between lines 90 and 107) to more explicitly underscore the distinctive aspects and scientific contributions of our work.

    Specifically, we have:

    1.  **Clarified the unique contribution of this paper** by emphasizing the broader geographical range and methodological scope of our drone remote sensing application for intertidal habitat mapping, as well as our focus on classifying diverse macrophyte types with high accuracy.

    2.  **Highlighted the novelty of our approach in integrating multiple spatial scales** through the simulation of satellite resolutions and quantification of the impact of spatial resolution on classification accuracy, providing a framework to better understand challenges of satellite based-habitat mapping.

    3.  **Added a summary of the study’s contributions** to contextualize our findings within existing research and underscore the broader implications of this work for coastal habitat mapping.

        We believe these additions better address the innovative aspects of our study and strengthen the scientific contribution of the manuscript.

2.  Thank you for your detailed feedback on the classification results presented in Figures 5 to 8. We appreciate your attention to the spectral representations of the Magnoliopsida category in the RGB images. Your observations regarding the varying appearances of pixels classified as seagrass in the RGB images were extremely helpful, as this distinction between different coloration of seagrasses was not sufficiently emphasized in the original manuscript.

    In response, we have revised the main text to clarify that the Magnoliopsida class includes pixels with varying spectral signatures due to differences in physiological states, such as leaf health and senescence. These variations are visible in the RGB images, where healthy seagrass leaves appear green, while senescent or altered leaves may appear brown or even white. This variation in seagrass leaf color has been observed during the field campaign by our team.

    To ensure accurate classification despite these spectral differences, we carefully included spectral signatures from both healthy and senescent seagrass states in the neural network classifier’s training dataset. This approach allowed the classifier to consistently recognize both states as part of the same seagrass class, ensuring that all instances of seagrass, regardless of coloration, are correctly classified as Magnoliopsida.

    These modification of the text can be found between lines 188-191 and also between lines 274 and 278.

### 2024-11-12 docs/_tex/index.tex
adding reviewer comments

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

@@

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\usepackage{bookmark}

@@

Drones (Unmanned Aerial Vehicles -- UAVs) can potentially fill the data
gaps left by satellite RS and \emph{in situ} measurements, due to their
ability to provide spatially-explicit observations at very high spatial
resolutions (pixel size from mm to cm) while capturing data at
multi-spectral resolution \citep{fairley2022drone, oh2017use}. The
versatility of drones allows for their application across a diverse
thematic range , from coastal zone management
\citep{adade2021, casella2020, angnuureng2022} to mapping species
distribution
\citep{joyce2023, tallam2023, Roca2022, Roman2021, Brunier2022Topographic, sousa2019blue}.
However, when applied to coastal habitat mapping, previous case studies
were mostly limited to a low number of flights over a single study site,
restricting the generalizability of their application over wider
geographical scales
\citep{Roman2021, collin2019improving, rossiter2020uav, Brunier2022Topographic}.
The present study aimed at analyzing the potential of multispectral
drone RS to map intertidal macrophytes over a diverse biogeographical
range, with a particular focus on discriminating Magnoliopsida and
Chlorophyceae (Seagrass and Green Algae, respectively). Nine drone
flights were performed over soft-bottom intertidal areas along the
Atlantic coastlines of two European countries (France and Portugal),
covering a wide range of habitats, from monospecific seagrass meadows to
meadows mixed with green, or red macroalgae. A deep learning algorithm
was trained and validated for macrophyte discrimination, emphasizing
applicability across diverse sites without a loss of prediction
accuracy. The classification maps obtained at a very high spatial
resolution with the drone were spatially degraded to satellite
resolutions, making it possible to assess the effect of spatial
resolution on classification accuracy, and provide insights for coastal
habitat mapping using satellite remote sensing.

@@

Seven study sites distributed between France and Portugal were selected
for their extensive intertidal seagrass beds. Two sites were located in
the Gulf of Morbihan, France (Figure~\ref{fig-map} A : 47.5791°N,
2.8018°W). This gulf covers an area of 115 km² and is only connected to
the sea through a 900 m wide channel. A total of 53 small islands are
scattered across the gulf leading to 250 km of shorelines. Patchy
seagrass meadows can be found on many of these islands. One of the sites
within the gulf was on one its islands (Arz) and the other was located
further south on a mainland beach area (Duer). The Gulf of Morbihan is a
Natura 2000 site and a Regional Protected Area due to its rich
biodiversity, including its seagrass meadows, and is also classified as
a RAMSAR site, which highlights its significance as a wetland of
international importance. Two other sites were located in Bourgneuf Bay,
France (Figure~\ref{fig-map} B : 46.9849°N, 2.1488°W) which is a 340 km²
semi-enclosed macrotidal bay, protected from waves by Noirmoutier
Island. Bourgneuf bay hosts a large intertidal seagrass meadow of about
6 km² \citep{ZOFFOLI2020112020}. Within this meadow, the sites observed
by drones (L'Epine and Barbatre) contained monospecific beds of
\emph{Nanozostera noltei} (dwarf eelgrass, syn. \emph{Zostera noltei})
with very little mixing with other macrophytes. Bourgneuf Bay is also
part of the Natura 2000 network and serves as a RAMSAR site due to its
critical habitat for migratory bird species and its extensive seagrass
meadows \citep{Zoffoli2022}. Three sites were surveyed in the Ria de
Aveiro Coastal Lagoon in Portugal (Figure~\ref{fig-map} C : 40.6887°N,
8.6810°W). The extent of this lagoon is \textasciitilde83 km² (at low
tide) with many narrow channels, large salt marshes and many mudflats
that uncover at low tide \citep{sousa2017blue}. It is connected to the
open sea through a single channel, with a tidal lag between the North
and the South of the lagoon. The southernmost site (Gafanha) is a
mudflat located in the Mira channel (one of the four main channels of
the lagoon) whereas the two other sites (Mataducos and Marinha
Lanzarote) were situated in the middle of the lagoon and only accessible
by boat. These Portuguese sites are characterized by a more diverse
intertidal vegetation, where patches of seagrass intermingle with red,
brown, and green macroalgae. The Aveiro Lagoon, like the other study
areas, is a Natura 2000 site and a RAMSAR wetland, recognized for its
rich mosaic of habitats and importance for biodiversity, including
migratory bird species and intertidal vegetation.

@@

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Low_res/Fig1_Map_Drone_Sites.png}

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs with ten spectral bands ranging
from the blue to the near infrared (NIR): 444, 475, 531, 560, 650, 668,
705, 717, 740 and 840 nm. To ensure consistent lighting conditions
across flight paths, the drone's trajectory was aligned to maintain a
solar azimuth angle of 90 degrees. An overlap of 70\% and 80\% (side and
front respectively) between each image was set for each flight. A
downwelling light sensor (DLS2) was used to acquire irradiance data
concomitantly with the camera measurements. Raw data were calibrated in
reflectance using a calibration panel reflective at \textasciitilde50\%
provided by the manufacturer. Across all sites, flights were made at two
different altitudes : 12 m or/and 120 m, with a spatial resolution of 8
mm and 80 mm, respectively (Table~\ref{tbl-flights}).

@@

\includegraphics[width=6.63in,height=\textheight]{Figures/High_res/Table1/table_flights.png}

@@

Before each flight, targets used as ground control points were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). Ground control points were used to
correct georeferencing imprecision of orthomosaics with an horizontal
and vertical accuracy of 10cm. A dGPS was also used to georeference
quadrats of 0.25 m² ,which assessed the presence or absence of five key
taxonomic classes of intertidal vegetation : Bacillariophyceae
(unicellular benthic diatoms forming biofilms at the sediment surface
during low tide), Phaeophyceae (brown macroalgae), Magnoliopsida (dwarf
eelgrass), Chlorophyceae (green macroalgae) and Rhodophyceae (red
macroalgae) (Figure~\ref{fig-vegetation}). Only homogeneous vegetation
patches extending over several meters were selected as ground control
points. Pictures of each quadrat were uploaded online to the open-portal
Global Biodiversity Information Facility (GBIF) platform
\citep{BedeGbif}. Each photograph was also processed to estimate the
percent cover of each type of vegetation using an image processing
software \citep[ImageJ,][]{schneider2012nih}. Hyperspectral reflectance
signatures of each vegetation class were recorded using an ASD FieldSpec
HandHeld 2 spectroradiometer, which acquires reflectance between 325 and
1075 nm, with 1 nm of spectral resolution. Hyperspectral signatures
served dual purposes: they validate the radiometric calibration of drone
data and contribute to misclassification reduction in photo
interpretations.

@@

\includegraphics[width=1\textwidth,height=\textheight]{Figures/Low_res/Spectral_shapes_total.png}

@@

\caption{\label{fig-vegetation}The five taxonomic classes of vegetation
used to train the Neural Network model and their raw spectral signatures
at the spectral resolution of the Micasense RedEdge Dual MX. A :
Magnoliopsida (Nanozostera noltei syn. Zostera noltei) ; B :
Phaeophyceae (Fucus sp.) ; C : Rhodophyceae (Gracilaria vermiculophylla)
; D : Chlorophyceae (Ulva sp.) ; E : Bacillariophyceae (Diatoms - MPB)}

@@

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Low_res/Figure3_workflow.png}

@@

\includegraphics[width=6.63in,height=\textheight]{Figures/High_res/table_validation_px.png}

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model. The training pixels were
categorized into seven different classes, representing the various
habitats encountered at the different study sites: Sediment, Water,
Chlorophyceae, Magnoliopsida, Bacillariophyceae, Phaeophyceae and
Rhodophyceae. Only data from the low-altitude flights
(Table~\ref{tbl-flights}) were used for training because their 8 mm
spatial resolution allowed to avoid spectral sub-pixel mixing and to
accurately identify vegetation classes. More than 418,000 pixels at 8 mm
resolution from the 3 training flights were used to train the model
(Table~\ref{tbl-validationPX}). Twenty one variables were used by the
model as predictors: the ten raw spectral bands of the Micasense RedEdge
Dual MX multispectral camera (ranging from 444 nm to 840 nm), the same
ten spectral bands standardized using a min/max transformation
(Equation~\ref{eq-std} ; \citep{Cao2017}) and the Normalized difference
vegetation index (NDVI, Equation~\ref{eq-ndvi}). Standardisation of
spectral bands is commonly used to eliminate the scaling differences
between spectra and to limit the effect of biomass on the spectra shape
\citep{Douay2022, Davies2023}.

@@

The classification model was applied to all flights at both 12 and 120 m
of altitude. \emph{In situ} information on georeferenced class type and
percent cover, acquired over homogeneous vegetation patches at the same
time as drone flights was used to assess the model accuracy. These
images were used to construct a validation dataset indicating the
presence or absence of each class. Additionally to the quadrat-based
validation dataset, polygons of each class were photo interpreted in
order to increase the number of pixels of the validation dataset. A
total of 536,000 pixels were used to validate the Neural Network
classifier. The sites with the lowest and highest number of validation
data were Gafanha Low (17316 pixels) and Marinha Lanzarote (159713
pixels), respectively. A confusion matrix, along with precision metrics
such as global accuracy, sensitivity, specificity, F1 score, and Kappa
coefficient, were generated for each site. All validation matrices were
then aggregated to create an overall matrix .

@@

To evaluate the influence of spatial resolution on the model's output,
we resampled the drone orthomosaics from their native resolution (8 cm
for the high altitude flights) using the ``average'' method from the
Terra package in R. DISCOV was then applied to these resampled rasters,
and the results were compared to the original model predictions. For
each resolution and vegetation class, we calculated the predicted area
loss, where a score of 0 indicates no area loss during spatial
resampling, and a score of 100 indicates complete loss of the vegetation
class.

We used a Generalized Linear Model (GLM) with a Beta distribution to
examine the relationship between pixel resolution, vegetation class, and
their interaction on the loss of vegetation. The loss of vegetation was
modelled as function of the interaction between pixel resolution and
vegetation class (Bacillariophyceae, Phaeophyceae, Magnoliopsida,
Chlorophyceae and Rhodophyceae). Sample vs fitted residuals and
quartile-quartile graphics were assessed visually, to ensure assumption
of the models used were met

@@

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/High_res/Compare_reflectance_both.png}

@@

Each drone flight was used to produce a prediction map, as well as a
probability map that indicates the model derived probability of the
selected class for every pixel. The low-altitude flight conducted in
Gafanha, Portugal, represented the site with the highest complexity
(Figure~\ref{fig-GafLow}). Among the five vegetation classes on which
the model was trained, four were present on this site, with green and
red macroalgae mixed with a seagrass meadow. There was also diatoms
forming biofilms on bare sediment surface. Although the seagrass was
solely composed of a single specie, \emph{Nanozostera noltei}, various
colors of this specie could be observed from dark green (corresponding
to healthy leaves) to brown (when leaves are senescent or have an
altered pigment composition). Regardless of the variation of color, the
class Magnoliopsida was accurately predicted by the model (F1 score of
0.96 at that site).

@@

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Low_res/Maps Pred/FigX-Gaf_Low_Pred.png}

@@

The high-altitude flight over Gafanha covered a total area of
\textasciitilde1 km² (Figure~\ref{fig-GafHigh}). A channel contouring a
small island was masked in the prediction map. Most of vegetation area
was classified as seagrass by the model, including patches with
discolored leaves. Only a few pixels were classified as green algae (F1
score of 0.55). Patches of Rhodophyceae were correctly classified (F1
score of 0.85). In the northern part of the site and near the land
eadges, patches of the schorre angiosperm \emph{Sporobolus maritimus}
(syn. \emph{Spartina maritima)} were misclassified, either as seagrass
or as brown algae (F1 score of 0.77 and 0.71, respectively).

@@

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Low_res/Maps Pred/FigX-Gaf_High_Pred1.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Low_res/Maps Pred/FigX-Boat_Pred.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Low_res/Maps Pred/FigX-Dike_Pred.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{./Figures/Low_res/Validation/ConfusionMatrixGlobal.png}

@@

\includegraphics[width=1\textwidth,height=\textheight]{Figures/Low_res/VIP/Fig_VIP.png}

@@

Clear differences were seen in vegetation loss across different
resolutions and vegetation classes (@fig-pixel). At a fine resolution of
1m, changes in the retrieved area for each vegetation type are minimal.
Green algae shows the highest loss, with 1.2\% area lost compared to the
native resolution (80 mm). As the resolution coarsens to 10m, vegetation
loss becomes more pronounced, with green algae again experiencing the
greatest reduction (12\% compared to 8cm) and seagrass showing the
smallest loss (1.3\%). At a resolution of 30m, all green algae have been
lost (100\% compared to 8cm), while seagrass experiences a relatively
small reduction of 11\%. Brown algae and red algae show moderate
declines, with losses at 30m resolution reaching approximately 37\% and
59\%, respectively.

@@

\includegraphics[width=0.9\textwidth,height=\textheight]{Figures/High_res/resolution_plot.png}

@@

\includegraphics[width=0.9\textwidth,height=\textheight]{Figures/Low_res/Upscaling/density_vs_Proportion.png}

>>>>

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

@@

\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\usepackage{bookmark}

@@

Drones (Unmanned Aerial Vehicles -- UAVs) can potentially fill the data
gaps left by satellite RS and \emph{in situ} measurements, due to their
ability to provide spatially-explicit observations at very high spatial
resolutions (pixel size from mm to cm) while capturing data at
multi-spectral resolution \citep{fairley2022drone, oh2017use}. The
versatility of drones allows for their application across a diverse
thematic range , from coastal zone management
\citep{adade2021, casella2020, angnuureng2022} to mapping species
distribution
\citep{joyce2023, tallam2023, Roca2022, Roman2021, Brunier2022Topographic, sousa2019blue}.
However, when applied to coastal habitat mapping, previous case studies
were mostly limited to a low number of \textbf{drone} flights over a
single study site, restricting the generalizability of their application
over wider geographical scales
\citep{Roman2021, collin2019improving, rossiter2020uav, Brunier2022Topographic}.
\textbf{These studies have demonstrated the capability of drones to map
intertidal habitats, including seagrasses; however, broader
generalization of these findings is still lacking. The current paper
uniquely expands the spatial and methodological scope of drone-based
remote sensing for intertidal habitat mapping across a broad
biogeographical range. It demonstrates the feasibility of accurately
classifying diverse macrophyte types across various study sites, with a
particular focus on distinguishing Magnoliopsida (seagrasses) and
Chlorophyceae (green algae). Unlike previous studies, our approach
integrates multiple spatial scales by simulating satellite resolutions
and quantifying the impact of spatial resolution on classification
accuracy.} Nine drone flights were performed over soft-bottom intertidal
areas along the Atlantic coastlines of two European countries (France
and Portugal), covering a wide range of habitats, from monospecific
seagrass meadows to meadows mixed with green, or red macroalgae. A deep
learning algorithm was trained and validated for macrophyte
discrimination, emphasizing applicability across diverse sites without a
loss of prediction accuracy. The classification maps obtained at a very
high spatial resolution with the drone were spatially degraded to
satellite resolutions, making it possible to assess the effect of
spatial resolution on classification accuracy, and provide insights for
coastal habitat mapping using satellite remote sensing. \textbf{This
study is, therefore, among the first to quantify the effects of spatial
resolution on the accuracy of drone-based macrophyte classification
across a wide geographical scale, providing a framework for to better
understand satellite-based classification challenges.}

@@

Seven study sites distributed between France and Portugal were selected
for their extensive intertidal seagrass beds. Two sites were located in
the Gulf of Morbihan, France (Figure~\ref{fig-map} A : 47.5791°N,
2.8018°W). This gulf covers an area of 115 km² and is only connected to
the sea through a 900 m wide channel. A total of 53 small islands are
scattered across the gulf leading to 250 km of shorelines. Patchy
seagrass meadows can be found on many of these islands. One of the sites
within the gulf was on one its islands (Arz) and the other was located
further south on a mainland beach area (Duer). The Gulf of Morbihan is a
Natura 2000 site and a Regional Protected Area due to its rich
biodiversity, including its seagrass meadows, and is also classified as
a RAMSAR site, which highlights its significance as a wetland of
international importance. Two other sites were located in Bourgneuf Bay,
France (Figure~\ref{fig-map} B : 46.9849°N, 2.1488°W) which is a 340 km²
semi-enclosed macrotidal bay, protected from waves by Noirmoutier
Island. Bourgneuf bay hosts a large intertidal seagrass meadow of about
6 km² \citep{ZOFFOLI2020112020}. Within this meadow, the sites observed
by drones (L'Epine and Barbatre) contained monospecific beds of
\emph{Zostera noltei} (dwarf eelgrass) with very little mixing with
other macrophytes. Bourgneuf Bay is also part of the Natura 2000 network
and serves as a RAMSAR site due to its critical habitat for migratory
bird species and its extensive seagrass meadows \citep{Zoffoli2022}.
Three sites were surveyed in the Ria de Aveiro Coastal Lagoon in
Portugal (Figure~\ref{fig-map} C : 40.6887°N, 8.6810°W). The extent of
this lagoon is \textasciitilde83 km² (at low tide) with many narrow
channels, large salt marshes and many mudflats that uncover at low tide
\citep{sousa2017blue}. It is connected to the open sea through a single
channel, with a tidal lag between the North and the South of the lagoon.
The southernmost site (Gafanha) is a mudflat located in the Mira channel
(one of the four main channels of the lagoon) whereas the two other
sites (Mataducos and Marinha Lanzarote) were situated in the middle of
the lagoon and only accessible by boat. These Portuguese sites are
characterized by a more diverse intertidal vegetation, where patches of
seagrass intermingle with red, brown, and green macroalgae. The Aveiro
Lagoon, like the other study areas, is a Natura 2000 site and a RAMSAR
wetland, recognized for its rich mosaic of habitats and importance for
biodiversity, including migratory bird species and intertidal
vegetation.

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./Figures/Low_res/Fig1_Map_Drone_Sites.png}

@@

At each location, a DJI Matrice 200 quadcopter drone equipped with a
Micasense RedEdge Dual MX multispectral camera was flown to take 1.2
million pixel reflectance photographs with ten spectral bands ranging
from the blue to the near infrared (NIR): 444, 475, 531, 560, 650, 668,
705, 717, 740 and 840 nm. To ensure consistent lighting conditions
across flight paths, the drone's trajectory was aligned to maintain a
solar azimuth angle of 90 degrees. An overlap of 70\% and 80\% (side and
front respectively) between each image was set for each flight. A
downwelling light sensor (DLS2) was used to acquire irradiance data
concomitantly with the camera measurements. Raw data were calibrated in
reflectance using a calibration panel reflective at \textasciitilde50\%
provided by the manufacturer. Across all sites, flights were made at two
different altitudes : 12 m or/and 120 m, with a spatial resolution of 8
mm and 80 mm, respectively (Table~\ref{tbl-flights}). \textbf{Low
altitude flights were used as training dataset for a neural network
model while high altitude flight were used for validation.}

@@

\includegraphics[width=6.63in,height=\textheight,keepaspectratio]{Figures/High_res/Table1/table_flights.png}

@@

Before each flight, targets used as ground control points were
distributed over the study site and georeferenced with a Trimble © Geo
XH 6000 differential GPS (dGPS). Ground control points were used to
correct georeferencing imprecision of orthomosaics with an horizontal
and vertical accuracy of 10cm. A dGPS was also used to georeference
quadrats of 0.25 m² ,which assessed the presence or absence of five key
taxonomic classes of intertidal vegetation : Bacillariophyceae
(unicellular benthic diatoms forming biofilms at the sediment surface
during low tide, \textbf{ranging in size from small patches (m²) to
entire mudflats (km²)}; henceforth: Diatoms), Phaeophyceae (henceforth:
brown macroalgae), Magnoliopsida (henceforth: Seagrasses), Chlorophyceae
(henceforth: Green macroalgae) and Rhodophyceae (henceforth: red
macroalgae) (Figure~\ref{fig-vegetation}). Only homogeneous vegetation
patches extending over several meters were selected as ground control
points. Pictures of each quadrat were uploaded online to the open-portal
Global Biodiversity Information Facility (GBIF) platform
\citep{BedeGbif}. Each photograph was also processed to estimate the
percent cover of each type of vegetation using an image processing
software \citep[ImageJ,][]{schneider2012nih}. Hyperspectral reflectance
signatures of each vegetation class were recorded using an ASD FieldSpec
HandHeld 2 spectroradiometer, which acquires reflectance between 325 and
1075 nm, with 1 nm of spectral resolution. Hyperspectral signatures
served dual purposes: they validate the radiometric calibration of drone
data and contribute to misclassification reduction in photo
interpretations.

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{Figures/Low_res/Spectral_shapes_total.png}

@@

\caption{\label{fig-vegetation}The five taxonomic classes of vegetation
used to train the Neural Network model and an example of their raw
spectral signatures at the spectral resolution of the Micasense RedEdge
Dual MX. A : Magnoliopsida (\emph{Zostera noltei}) ; B : Phaeophyceae
(\emph{Fucus sp.}) ; C : Rhodophyceae (\emph{Gracilaria
vermiculophylla}) ; D : Chlorophyceae (\emph{Ulva sp.}) ; E :
Bacillariophyceae (Diatoms - MPB). \textbf{Classes were identified
following the World register of Marine species.}}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./Figures/Low_res/Figure3_workflow.png}

@@

\includegraphics[width=6.63in,height=\textheight,keepaspectratio]{Figures/High_res/table_validation_px.png}

@@

A dataset containing photo-interpreted drone reflectance pixels was
built to train a Neural Network model. The training pixels were
categorized into seven different classes, representing the various
habitats encountered at the different study sites: Sediment, Water,
green macroalgae, seagrasses, diatoms, brown macroalgae and red
macroalgae. Only data from the low-altitude flights
(Table~\ref{tbl-flights}) were used for training because their 8 mm
spatial resolution allowed to avoid spectral sub-pixel mixing and to
accurately identify vegetation classes. \textbf{In the field, seagrasses
displaying two distinct colors, each with a unique spectral signature,
were observed. Careful attention was given to incorporating training
pixels from both color types into the training dataset for the seagrass
class. This approach was consistently applied to all classes within the
model.} More than 418,000 pixels at 8 mm resolution from the 3 training
flights were used to train the model (Table~\ref{tbl-validationPX}).
Twenty one variables were used by the model as predictors: the ten raw
spectral bands of the Micasense RedEdge Dual MX multispectral camera
(ranging from 444 nm to 840 nm), the same ten spectral bands
standardized using a min/max transformation (Equation~\ref{eq-std} ;
\citep{Cao2017}) and the Normalized difference vegetation index (NDVI,
Equation~\ref{eq-ndvi}). Standardisation of spectral bands is commonly
used to eliminate the scaling differences between spectra and to limit
the effect of biomass on the spectra shape
\citep{Douay2022, Davies2023}.

@@

The classification model was applied to all flights at both 12 and 120 m
of altitude. \emph{In situ} information on georeferenced class type and
percent cover, acquired over homogeneous vegetation patches at the same
time as drone flights was used to assess the model accuracy. These
images were used to construct a validation dataset indicating the
presence or absence of each class. Additionally to the quadrat-based
validation dataset, polygons of each class were photo interpreted in
order to increase the number of pixels of the validation dataset. A
total of 536,000 pixels were used to validate the Neural Network
classifier. The sites with the lowest and highest number of validation
data were Gafanha Low (17316 pixels) and Marinha Lanzarote (159713
pixels), respectively. A confusion matrix, along with precision metrics
such as global accuracy, sensitivity, specificity, F1 score, and Kappa
coefficient, were generated for each site. These metrics were computed
as follow :

\[
\text{Global accuracy} = \frac{\sum_{i=1}^{k} \text{TP}_i}{\sum_{i=1}^{k} \left(\text{TP}_i + \text{FP}_i + \text{FN}_i \right)}
\] \[
\text{Sensitivity}_i = \frac{\text{TP}_i}{\text{TP}_i + \text{FN}_i}
\] \[
\text{Specificity}_i = \frac{\text{TN}_i}{\text{TN}_i + \text{FP}_i}
\] \[
\text{F1}_i = \frac{2 \cdot \text{TP}_i}{2 \cdot \text{TP}_i + \text{FP}_i + \text{FN}_i}
\] Where \(\text{TP}_i\), \(\text{TN}_i\), \(\text{FN}_i\) and
\(\text{FP}_i\) represent the true positives, true negatives, false
negatives and false positives relative to the class i.

All validation matrices were then aggregated to create an overall matrix

@@

To assess the impact of spatial resolution on the model's output, we
resampled the drone orthomosaics from their native resolution (8 cm for
high-altitude flights) using the ``average'' method from the terra
package in R. \textbf{The rasters were resampled to 32 different
resolutions, ranging from 10 cm to 30 m.} DISCOV was then applied to
these resampled rasters, and the results were compared to the original
model predictions. For each resolution and vegetation class, we
calculated the predicted area loss, where a score of 0 indicates no area
loss during spatial resampling, and a score of 100 indicates complete
loss of the vegetation class.

We used a Generalized Linear Model (GLM) with a Beta distribution to
examine the relationship between pixel resolution, vegetation class, and
their interaction on the loss of vegetation. The loss of vegetation was
modelled as function of the interaction between pixel resolution and
vegetation class (Diatoms, brown macroalgae, seagrass, green macroalgae
and red macroalage). Sample vs fitted residuals and quartile-quartile
graphics were assessed visually, to ensure assumption of the models used
were met.

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./Figures/High_res/Compare_reflectance_both.png}

@@

Each drone flight was used to produce a prediction map, as well as a
probability map that indicates the model derived probability of the
selected class for every pixel. The low-altitude flight conducted in
Gafanha, Portugal, represented the site with the highest complexity
(Figure~\ref{fig-GafLow}). Among the five vegetation classes on which
the model was trained, four were present on this site, with green and
red macroalgae mixed with a seagrass meadow. There was also diatoms
forming biofilms on bare sediment surface. Although the seagrass was
solely composed of a single specie, \emph{Zostera noltei}, various
colors of this specie could be observed from dark green (corresponding
to healthy leaves) to brown (when leaves are senescent or have an
altered pigment composition). Regardless of the variation of color, the
class Magnoliopsida (seagrass) was accurately predicted by the model (F1
score of 0.96 at that site).

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./Figures/Low_res/Maps Pred/FigX-Gaf_Low_Pred.png}

@@

The high-altitude flight over Gafanha covered a total area of
\textasciitilde1 km² (Figure~\ref{fig-GafHigh}). A channel contouring a
small island was masked in the prediction map. Most of vegetation area
was classified as seagrass by the model, including patches with
discolored leaves. Only a few pixels were classified as green macroalgae
(F1 score of 0.55). Patches of red macroalgae were correctly classified
(F1 score of 0.85). In the northern part of the site and near the land
eadges, patches of the schorre angiosperm \emph{Sporobolus maritimus}
(syn. \emph{Spartina maritima)} were misclassified, either as seagrass
or as brown algae (F1 score of 0.77 and 0.71, respectively).

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./Figures/Low_res/Maps Pred/FigX-Gaf_High_Pred1.png}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./Figures/Low_res/Maps Pred/FigX-Boat_Pred.png}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./Figures/Low_res/Maps Pred/FigX-Dike_Pred.png}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{./Figures/Low_res/Validation/ConfusionMatrixGlobal.png}

@@

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{Figures/Low_res/VIP/Fig_VIP.png}

@@

Clear differences were seen in vegetation loss across different
resolutions and vegetation classes (Figure~\ref{fig-pixelsize}). At a
fine resolution of 1m, changes in the retrieved area for each vegetation
type are minimal. Green algae shows the highest loss, with 1.2\% area
lost compared to the native resolution (80 mm). As the resolution
coarsens to 10m, vegetation loss becomes more pronounced, with green
macroalgae again experiencing the greatest reduction (12\% compared to
8cm) and seagrass showing the smallest loss (1.3\%). At a resolution of
30m, all green algae have been lost (100\% compared to 8cm), while
seagrass experiences a relatively small reduction of 11\%. Brown
macroalgae and red macroalgae show moderate declines, with losses at 30m
resolution reaching approximately 37\% and 59\%, respectively.

@@

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{Figures/High_res/resolution_plot.png}

@@

\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{Figures/Low_res/Upscaling/density_vs_Proportion.png}
